\sys: Light-touch Data-Parallel Shell Processing

### 2020-05-22 pldi/draft.tex
Section 3 mini-pass

\heading{Providing Combinators}
For most commands, the annotations are enough to allow parallelization: 
  commands either fall under \sta, having straightforward combiners, or fall under a subset of \pur that can be parallelized using divide-and-conquer techniques~\sx{optimizer}.
Unfortunately, there are commands that require less obvious tactics---after all, \unix commands are Turing-complete programs and failing to plan for this generality would inhibit \sys's usefulness.

To support the parallelization of arbitrary commands in \pur, \sys allows supplying custom \emph{map} and \emph{fold} functions.
In line with the \unix philosophy, these functions can be written in any language as long as they conform to a few invariants:
  (i) \emph{map} is in \sta and \emph{fold} is in \pur,
  (ii) \emph{map} can consume (or extend) the output of the original command and  \emph{fold} can consume (and combine) the results of multiple \emph{map} invocations, and
  (iii) their composition produces the same output with the original command.
If these invariants are met, after DFG analysis~\sx{ir} \sys can replace occurrences of these commands with their \emph{map} and \emph{fold} equivalents, enabling their parallelization.

@@

The core of \sys is a dataflow graph (DFG) model that is expressive enough to support a large subset of shell programs.\nv{Immediate Q: How Large? Which programs are not supported? Why?}
A fundamental difference with other DFG models is that it contains information about the order in which a node reads its inputs.
This enables a set of graph transformations that expose data parallelism as task parallelism.\nv{What does this mean? What if they are fundamentally different in the reivewer's mind? Example?}
These transformations can then be applied in an iterative way, optimizing the program to utilize available computational resources.

@@

The two main shell abstractions are
  (i) files (or pipes, \emph{viz.} correspondence between files and pipes in \S\ref{bg:pipelines}\nv{Not in this version. Should we call them "streams"?}) that contain data, and
  (ii) commands that communicate through these files.
\nv{quick solution: change all files to streams and make the equivalence explicit here}
\sys's DFG model represents commands as nodes and \todo{streams} as edges.

@@

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative fold $f'$, satisfying the following equation:

\[
\forall x, x', s, f(x \cdot x', s) = f'(m(x, s) \cdot m(x', s), s)
\]

\noindent
This means that we get the same output by applying $f$ to a
concatenation of two inputs $x, x'$ as by applying the fold function
$f'$ to the concatenation of the outputs produced by applying map $m$
to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

This transformation can extended straightforwardly to pure commands implemented by a map-fold pair
$(m, r)$ as $ v(x_1 \cdot x_2 \cdots x_n, s) \Rightarrow r(m(x_1, s)
\cdot m(x_2, s) \cdots m(x_n, s), s)$.
% \todo{It is shown in Fig.}\ref{fig:parallelization-transformation}.

>>>>

\heading{Providing \todo{Aggregators}}
For commands in \sta, the annotations are enough to enable parallelization: they are applied to partitions of their input in parallel, and their outputs are simply concatenated.
%% commands either fall under \sta, having straightforward aggregators, or fall under a subset of \pur that can be parallelized using divide-and-conquer techniques~\sx{optimizer}. \kk{pure commands are what we have to provide aggregators for, so this is not totally valid.}
%% Unfortunately, there are commands that
%% require less obvious tactics---after all, \unix commands are Turing-complete programs and failing to plan for this generality would inhibit \sys's usefulness. \kk{I would propose deleting this last sentence. we have already separated commands into classes, so this sentence makes it seem like the classes don't help at all.}
%
To support the parallelization of arbitrary commands in \pur, \sys allows supplying custom \emph{map} and \emph{aggregate} functions.
In line with the \unix philosophy, these functions can be written in any language as long as they conform to a few invariants:
  (i) \emph{map} is in \sta and \emph{aggregate} is in \pur,
  (ii) \emph{map} can consume (or extend) the output of the original command and  \emph{aggregate} can consume (and combine) the results of multiple \emph{map} invocations, and
  (iii) their composition produces the same output with the original command.
If these invariants are met, after DFG analysis~\sx{ir} \sys can replace occurrences of these commands with their \emph{map} and \emph{aggregate} equivalents, enabling their parallelization.

@@

The core of \sys is an abstract dataflow graph (DFG) model that is used as the intermediate representation on which \sys performs optimizations.
%% representation that is expressive enough to model data parallelism in shell programs.
%% support a large subset of shell programs.\nv{Immediate Q: How Large? Which programs are not supported? Why?}
\sys first lifts sections of the input script to the DFG
representation, it performs transformations to expose parallelism, and
then instantiates each DFG back to a parallel shell script.  A
fundamental difference with other DFG models is that it encodes the
order in which a node reads its inputs, which in turn enables a set of
graph transformations that can be iteratively applied to expose
parallelization opportunities for \tsta and \tpur commands.
%% data parallelism as task parallelism.\nv{What does this mean? What if they are fundamentally different in the reivewer's mind? Example?}
%% These transformations can then be applied in an iterative way, optimizing the program to utilize available computational resources. 
In \Cref{graph-components} we define the DFG model and describe the
correspondence with shell scripts, and in \Cref{ir:transformations} we
describe the graph transformations and \todo{give intuition about their
correctness}.

@@

The two main shell abstractions are
  (i) streams, i.e. files or pipes, that contain data, and
  (ii) commands that communicate through these streams.
%% \nv{quick solution: change all files to streams and make the equivalence explicit here}
\sys's DFG model represents commands as nodes and streams as edges.

@@

Similarly, some pure commands (such as \ttt{sort} and \ttt{wc}) can be
parallelized using divide-and-conquer techniques. More formally, these
pure commands $f$ can be implemented as a combination of a function map
$m$ and an associative aggregate $agg$, satisfying the following equation:

\[
\forall x, x', s, f(x \cdot x', s) = agg(m(x, s) \cdot m(x', s), s)
\]

\noindent
This means that we get the same output by applying $f$ to a
concatenation of two inputs $x, x'$ as by applying the aggregation function
$agg$ to the concatenation of the outputs produced by applying map $m$
to each of $x$ and $x'$.
%% \km{We had also discussed here the case of
%%   operators that can be described as associative and commutative
%%   functions on the data (e.g., sum). In this case the parallelization
%%   has the flavor of data parallelism: one can split the input stream
%%   in an arbitrary way (e.g., randomly or using some other
%%   load-balancing scheme) and merge the outputs. For the more broad
%%   class of associative operations (e.g., sort), the formulation of
%%   this paragraph makes perfect sense. I just want to mention that this
%%   feels a lot more like \textbf{batching} than map-reduce-like data
%%   parallelization.} \kk{Thanks, I think you are right. Batching is a
%%   better name for it. In the end we didn't handle
%%   commutative-associative function (that are more like map-reduce) and
%%   so we changes the name to map-fold to not imply commutativity.}

@@

This transformation can extended straightforwardly to pure commands
implemented by a map-aggregate pair $(m, agg)$ as $ v(x_1 \cdot x_2
\cdots x_n, s) \Rightarrow agg(m(x_1, s) \cdot m(x_2, s) \cdots m(x_n,
s), s)$.
% \todo{It is shown in Fig.}\ref{fig:parallelization-transformation}.
